\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}
\tikzset{ ->, % makes the edges directed
>=stealth', % makes the arrow heads bold
node distance=2cm, % specifies the minimum distance between two nodes. Change if
necessary.
every state/.style={thick, fill=gray!10},
% sets the properties for each ’state’ node
initial text=$ $ }

\title{Tarea 01}
\author{Ana Maritza Bello Yañez}
\date{\today}

\begin{document}
\maketitle

\begin{enumerate}
\item The following are the state diagrams of two DFAs, $M_1$ and $M_2$ . Answer the
following questions about each of these machines.


\begin{figure}[h!]
\centering
    \begin{tikzpicture}
        \node [state, initial] (q1){$q_1$}; 
        \node [state, accepting,right of = q1] (q2){$q_2$};
        \node [state, below of = q1, xshift=1cm] (q3){$q_3$};
            
        \draw (q1) edge[loop above] node{$b$} (q1);
        \draw (q1) edge[bend left, above] node{$a$} (q2);
        \draw (q3) edge[bend left, below] node{$b$} (q1);
        \draw (q2) edge[bend left, below] node{$a,b$}(q3);
        \draw (q3) edge[bend left, above] node{$a$} (q2);
    \end{tikzpicture}
\caption{$M_1$}
\end{figure}
    

        \begin{itemize}
\item What is the start state?
\item What is the set of accept states?
\item What sequence of states does the machine go through on input \texttt{aabb}
\item Does the machine accept the string \texttt{aabb}
\item Does the machine accept the string $\epsilon$?
        \end{itemize}

\begin{figure}[h!]
\centering
    \begin{tikzpicture}
        \node [state, initial, accepting] (q1){$q_1$}; 
        \node [state, right of = q1] (q2){$q_2$};
        \node [state, below of = q1] (q3){$q_3$};
        \node [state, accepting, right of = q3] (q4){$q_4$};
            
        \draw (q1) edge[loop above] node{$a$} (q1);
        \draw (q4) edge[loop right] node{$b$} (q4);
        \draw (q1) edge[bend left, above] node{$b$} (q2);
        \draw (q2) edge[bend left] node{$\ \ b$} (q4);
        \draw (q2) edge[bend left, below] node{$a$}(q3);
        \draw (q3) edge[bend left, above] node{$a$} (q2);
        \draw (q3) edge[bend left] node{$b\ \ $} (q1);
        \draw (q4) edge[bend left, below] node{$a$} (q3);
    \end{tikzpicture}
\caption{$M_2$}
\end{figure}

        \begin{itemize}
\item What is the start state?
\item What is the set of accept states?
\item What sequence of states does the machine go through on input \texttt{aabb}
\item Does the machine accept the string \texttt{aabb}?
\item Does the machine accept the string $\epsilon$?
        \end{itemize}

\item Give the formal description of the machines M1 and M2 pictured in Exercise
1.


\item The formal description of a DFA $M$ is
($\{q_1,q_2,q_3,q_4,q_5\}$,$\{u,d\}$,$\delta$,$q_3$,$\{q_3\}$), where $\delta$
is given by the following table. Give the state diagram of this machine.

\begin{center}
\begin{tabular}{c|cc}
            &   \texttt{u}     &   \texttt{d} \\
    \hline
    $q_1$   &   $q_1$   &   $q_2$   \\
    $q_2$   &   $q_1$   &   $q_3$   \\
    $q_3$   &   $q_2$   &   $q_4$   \\
    $q_4$   &   $q_3$   &   $q_5$   \\
    $q_5$   &   $q_4$   &   $q_5$
\end{tabular}
\end{center}


\item Each of the following languages is the intersection of two simpler
languages. In each part, construct DFAs for the simpler languages, then combine
them using the construction discussed in footnote 3 (page 46) to give the state
diagram of a DFA for the language given. In all parts $\sum = \{\texttt{a}, \texttt{b}\}$.

    \begin{enumerate}
        \item \{$w|w$ en has at least three \texttt{a}'s and at least two \texttt{b}'s\}
        \item \{$w|w$ has at exactly two \texttt{a}'s and at least two \texttt{b}'s\}
        \item \{$w|w$ has an even number of \texttt{a}'s and one or two \texttt{b}'s\}
        \item \{$w|w$ has an even number of \texttt{a}'s and each a is followed by at least one \texttt{b}\}
        \item \{$w|w$ has an even number of \texttt{a}'s and one or two \texttt{b}'s\}
        \item \{$w|w$ in has an odd number of \texttt{a}'s and ends with a \texttt{b}\}
        \item \{$w|w$ has even length and an odd number of \texttt{a}'s\}
    \end{enumerate}

\item Each of the following languages is the complement of a simpler language.
In each part, construct a DFA for the simpler language, then use it to give the
state diagram of a DFA for the language given. In all parts $\sum =
\{\texttt{a}, \texttt{b}\}$.

\begin{enumerate}
\item \{$w|w$ does not contain the substring \texttt{ab}\}
\item \{$w|w$ does not contain the substring \texttt{baba}\}
\item \{$w|w$ contains neither the substrings \texttt{ab} nor \texttt{ba}\}
\item \{$w|w$ is any string not in $\texttt{a}^* \texttt{b}^*$\}
\item \{$w|w$ is any string not in \texttt{(ab$^+$)$^*$}\}
\item \{$w|w$ is any string not in \texttt{a$^*\cup$b$^*$}\}
\item \{$w|w$ is any string that doesn't contain exactly two \texttt{a}'s\}
\item \{$w|w$ is any string except \texttt{a} and \texttt{b}\}
\end{enumerate}

\item Give state diagrams of DFAs recognizing the following languages. In all
parts the alphabet is {0,1}.

\begin{enumerate}
\item \{$w|w$ begins with a \texttt{1} and ends with a \texttt{0}\}
\item \{$w|w$ contains at least three \texttt{1}s\}
\item \{$w|w$ contains the substring \texttt{0101}, i.e., $w = x\texttt{0101}y$
for some $x$ and $y$\}
\item \{$w|w$ has length at least 3 and its third symbol is a \texttt{0}\}
\item \{$w|w$ starts with \texttt{0} and has odd length, or starts with
\texttt{1} and has even length\}
\item \{$w|w$ doesn't contain the substring \texttt{110}\}
\item \{$w|w$ the length of $w$ is at most 5\}
\item \{$w|w$ is any string except \texttt{11} and \texttt{111}\}
\item \{$w|w$ every odd position of $w$ is a \texttt{1}\}
\item \{$w|w$ contains at least two Os and at most one \texttt{1}\}
\item \{$\epsilon$, $0$\}
\item \{$w|w$ contains an even number of \texttt{O}s, or contains exactly two
\texttt{1}s\}
\item The empty set
\item All strings except the empty string
\end{enumerate}
    
\end{enumerate}

\end{document}
